Finding Quality Issues in SKOS Vocabularies

### 2012-03-28 background.tex
merged related work into 2

\section{Background}\label{sec:background}

% About Quality in SKOS vocabularies

@@

%In the course of the LATC project, Gueret et al. propose an infrastructure that utilizes crowd-sourcing technology to semi-automatically create links between datasets of the LOD cloud. To ensure the quality of the resulting network, human judgement is supported by calculating a report of locally approximated network measures. \todo{BH}{ich denke antoine hat das paper gemeint: http://www.mendeley.com/download/public/18928/4084690763/56bdde328dfbba2cb8594b57152c0e132ef00a40/dl.pdf . Leider fehlen da etliche Details um sich ein konkretes Bild machen zu können}

@@

% Argue why we need criteria that go beyond low-level integrity constraints and why we need to formalize them

Typical application areas of controlled vocabularies are classification, indexing, autocompletion, query reformulation/expansion, or serving as a glossary. As we discussed in detail in our earlier work \cite{Nagy2011}, these application areas have specific requirements with respect to vocabulary features, such as structure, availability, and documentation of conceptual resources. A vocabulary that doesn't fulfill these requirements can have a negative effect on an application's output and would therefore be judged by the application designers and potentially also by users as being a low-quality vocabulary. Missing relations between conceptual resources, for instance, can reduce retrieval recall when a vocabulary is used for query expansion. Missing entry points to concepts in a vocabulary can impede orientation for human users.

% Argue why we need to formalize them -> because otherwise they will be implemented inconsistently across applications
Thus, in order to help vocabulary and application designers in judging and improving the quality of SKOS vocabularies we must identify possible quality issues that go beyond integrity conditions defining data model consistency. We can identify such issues by examining quality criteria for controlled vocabularies that have already been discussed in standardized guidelines~\cite{ISO25964-1:2011,Z39.19:2005}, manuals~\cite{Svenonius2003,Hedden2010,Aitchison2000,Harpring2010}, tutorials~\cite{Soergel2002}, and scholarly articles \cite{Coronado2009,Soergel1997,Elkin2002,Kless2010}. Since we want to provide tool support for quality checking, we must formalize the identified issues into computable quality checking functions taking into account the syntax, structure and semantics of SKOS. This also ensures consistent implementation across tools.

>>>>

\section{Background and Related Work}\label{sec:background}

% Data quality in general
The problem of ``vocabulary quality'' is closely related to the more general problem of ``data quality`'' and has been discussed in data and information systems research (cf.~\cite{Batini2009}). Pipino et al.~\cite{Pipino2002} argue that dealing with data quality should involve both ``subjective perceptions of the individuals'' and ``objective measurements based on the data set in question''. We see our work as a contribution to the latter.

\subsection{Quality in SKOS Vocabularies}

@@

%In the course of the LATC project, Gueret et al. propose an infrastructure that utilizes crowd-sourcing technology to semi-automatically create links between datasets of the LOD cloud. To ensure the quality of the resulting network, human judgement is supported by calculating a report of locally approximated network measures.

@@

\subsection{Quality in Controlled Vocabularies}

Typical application areas of controlled vocabularies are classification, indexing, autocompletion, query reformulation/expansion, or serving as a glossary. As we discussed in detail in our earlier work \cite{Nagy2011}, these areas impose specific requirements on vocabulary features, such as structure, availability, and documentation of conceptual resources. Quality aspects of controlled vocabularies have already been discussed in standardized guidelines~\cite{ISO25964-1:2011,Z39.19:2005}, manuals~\cite{Svenonius2003,Hedden2010,Aitchison2000,Harpring2010}, tutorials~\cite{Soergel2002}, and scholarly articles \cite{Coronado2009,Soergel1997,Elkin2002,Kless2010}. Our work builds on this literature and identifies possible quality issues that go beyond model-level integrity conditions.

\subsection{Quality in Semantic Web and Linked Data Research}

Hogan et. al~\cite{Hogan2010} identify four categories of common errors and shortcomings in RDF documents and also Heath and Bizer~\cite{Heath2011} summarize best practices for publishing data on the Web. Also ontology evaluation, which is the task of measuring the quality of an ontology, has been discussed by Vrandecic~\cite{Vrandecic2010}. However, the authors focus on RDF datasets and ontologies in general without considering SKOS-specific properties.

### 2012-03-28 criteria.tex
merged related work into 2

\paragraph{\textbf{Definition (SKOS Vocabulary).}}
Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV, CS \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C \subseteq IR\) with \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) \todo{CM}{check exponent}being the set of \textbf{concepts}, \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating concepts with one another, $LV \subseteq I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of untyped \textbf{plain literals}, and \(CS I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I})\) being the set of \textbf{concept schemes}. Further, we let $V$ be the fully entailed RDFS interpretation of the underlying RDF graph, enriched by entailment of \texttt{owl:inverseOf} properties as well as subproperties of \texttt{owl:TransitiveProperty} and \texttt{owl:SymmetricProperty}.

@@

In the following, we explain the origins and design rationale for each quality issue and define the corresponding quality checking function. For better readability and due to lack of space we provide only semi-formal definitions and refer to the \sout{online documentation and the} implementation source code\footnote{\url{https://github.com/cmader/qSKOS/}} for further details.

>>>>

\todo{CM}{use definition environment and find out how to turn of numbering}
\paragraph{\textbf{Definition (SKOS Vocabulary).}}
Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV, CS \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C \subseteq IR\) with \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) \todo{CM}{check exponent}being the set of \textbf{concepts}, \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating concepts with one another, $LV \subseteq I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of untyped \textbf{plain literals}, and \(CS = I_{CEXT}(\texttt{skos:ConceptScheme}^\mathcal{I})\) being the set of \textbf{concept schemes}. Further, we let $V$ be the fully entailed RDFS interpretation of the underlying RDF graph, enriched by entailment of \texttt{owl:inverseOf} properties as well as subproperties of \texttt{owl:TransitiveProperty} and \texttt{owl:SymmetricProperty}.

@@

In the following, we explain the origins and design rationale for each quality issue and define the corresponding quality checking function. For better readability and due to lack of space we provide only semi-formal definitions and refer to the online documentation and the implementation source code\footnote{\url{https://github.com/cmader/qSKOS/}} for further details.

### 2012-03-28 main.tex
incorporated comments by antoine
merged related work into 2

\usepackage{color}

@@

% metrics = quantitative indicator of sth

\title{Finding Quality Issues in SKOS Vocabularies}
\author{Christian Mader\inst{1} \and Bernhard Haslhofer\inst{2}}
\institute{
	University of Vienna, Faculty of Computer Science\\\email{christian.mader@univie.ac.at}
	\and Cornell University, Information Science\\\email{bernhard.haslhofer@cornell.edu}}

@@

\input{related_work}

>>>>

\usepackage{color}
%\usepackage{ulem} Why this?

@@

%Title ideas:
% - How can we fix quality issues in SKOS vocabularies?
% - Detecting quality issues in SKOS Vocabularies
% - Computable Quality Guidlines for SKOSO Vocabularies
% - ...

\title{Finding Quality Issues in SKOS Vocabularies}
\subtitle{\todo{ALL}{I am still not happy with the title; better ideas?}}
\author{Christian Mader\inst{1} \and Bernhard Haslhofer\inst{2} \and Antoine Isaac\inst{3}}
\institute{
	University of Vienna, Faculty of Computer Science, Austria\\\email{christian.mader@univie.ac.at}
	\and Cornell University, Department of Information Science, USA\\\email{bernhard.haslhofer@cornell.edu}
	\and Vrije Universiteit Amsterdam, The Netherlands\\\email{aisaac@few.vu.nl}
	}

@@
